### 2011-04-07 main_allen.tex
Partially done merging Allen's changes

\documentclass[10pt,twocolumn,letterpaper]{article}
% ICCV Rules:
% 8 page limit (6 free + 2 paid)
% 10MB pdf file size limit
% 30MB supplementary materials file size limit (PDF or ZIP)

\usepackage{cvpr}
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{algorithm}
\usepackage{algorithmic}

% Needed for getting formulas in dia diagrams.
\usepackage{pstricks}
\usepackage{tikz}

\usepackage{mydefs}

% \cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{257} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi
\begin{document}

%\title{Accelerating $\ell_1$-Minimization Using Many-Core CPUs/GPUs \\ and Application to Face Recognition
%\title{Efficient Parallelization of Sparse Representation for Face Recognition}
%\thanks{Corresponding author: . This work was partially supported by ARO MURI W911NF-06-1-0076.}}
%\title{Many-Core Parallelization of $\ell_1$-Minimization for Face Recognition
\title{Parallelization of Fast $\ell_1$-Minimization for Face Recognition}

\author{First Author\\
Institution1\\
Institution1 address\\
{\tt\small firstauthor@i1.org}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Second Author\\
Institution2\\
First line of institution2 address\\
{\small\url{http://www.author.org/~second}}
}

\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction} 
$\ell_1$-minimization ($\ell_1$-min) has received much attention in recent
years due to important applications in compressive sensing
\cite{BrucksteinA2007} and sparse representation \cite{WrightJ2010-PIEEE}.  
$\ell_1$-min refers to finding the minimum $\ell_1$-norm solution to an
underdetermined linear system $\bb=A\xx$:
\begin{equation}
\min \|\xx\|_1\quad \mbox{ subj. to }\quad \bb = A\xx.
\label{eq:l1min}
\end{equation}
It is now well established that, under certain conditions
\cite{CandesE2005-IT_1,DonohoD2004}, the minimum $\ell_1$-norm solution is also
the \emph{sparsest} solution to the system \eqref{eq:l1min}.

Among its many applications, $\ell_1$-min has been recently used to reformulate
image-based face recognition as a sparse representation problem
\cite{WrightJ2009-PAMI}.  If we stack the training images of $K$ subject
classes $(A_1\in\Re^{d\times n_1}, \cdots, A_K\in\Re^{d\times n_K})$ into the
columns of a matrix $A = [A_1, \cdots, A_K]\in\Re^{d\times n}$, given a new
query image in the same vector form $\bb\in\Re^d$, \emph{sparsity-based
classification} (SBC) solves the following minimization problem:
\begin{equation}
\min_{\xx, \ee} \| \xx \|_1 + \|\ee\|_1 \quad \subj \quad \bb = A \xx + \ee.
\label{eq:l1min_denoise}
\end{equation}
If the sparsest solutions for $\x$ and $\e$ are recovered, $\ee$ provides a
means to compensate for pixels that are corrupted due to occlusion of some part of the query
image, and the dominant nonzero coefficients of $\xx$ reveal the membership of
$\bb$ based on the training image labels associated with $A$.  Since $A$ ideally
contains many images per user taken under different illuminations, 
$A\x$ also acts as a linear illumination model for the test image $\bb$.

In this paper, we study parallelization of $\ell$-min on many-core CPU and GPU
architectures. Although $\ell_1$-min \eqref{eq:l1min} is a convex
program, conventional algorithms such as interior-point methods
\cite{ChenS2001-SIAM,TibshiraniR1996} are known to suffer from poor scalability
for large-scale applications such as face recognition. Recently, a number of
accelerated algorithms have been proposed that explicitly take advantage of
the special structure of $\ell_1$-min problems
\cite{LorisI2009,YangA2010-ICIP}. We investigate parallelization of a
state-of-the-art $\ell_1$-min solution based on the classical framework of
\emph{augmented Lagrangian methods} (ALM) \cite{BertsekasD2003,YangA2010-ICIP}.

While solving \eqref{eq:l1min_denoise} can achieve exceptional
recognition accuracy on public databases, its success also relies on the assumption
that the query image is well aligned with the training images. In
\cite{WagnerA2009-CVPR}, the framework of SBC was extended to iteratively align
a query image to one subject class. The face alignment algorithm in turn solves
the following minimization problem:
\begin{equation}
\hat{\tau}_i = \arg\min_{\xx, \ee, \tau_i} \|\ee\|_1\quad \mbox{subj. to}\quad \tilde\bb\circ\tau_i = A_i\xx + \ee,
\label{eq:l1min_alignment}
\end{equation}
where $\tau_i\in T$ is in a finite-dimensional group $T$ of transformations
(e.g., affine or homography) acting on the image domain, which aligns $\tilde\bb$
with the training (frontal) images from the $i$-th class $A_i$. The algorithm 
extends Lucas-Kanade iterative alignment \cite{LucasB1981} using the
$\ell_1$-norm as a robust error function.

Numerically, the original solution \cite{WagnerA2009-CVPR} was implemented based 
on standard interior-point methods, and its speed has been since improved by ALM in a more recent paper
\cite{WagnerA2011-PAMI}. However, due the high per-class cost
of the alignment step, the recognition system still fails to achieve \emph{real-time recognition} against datasets of hundreds or thousands of users.
In this paper, in addition to accelerating the generic
$\ell_1$-min objectives \eqref{eq:l1min} and \eqref{eq:l1min_denoise}, we 
also discuss how to properly accelerate the iterative face alignment step
\eqref{eq:l1min_alignment} on multi-core CPUs and GPUs.

In this paper we will clearly distinguish between algorithm {\em parallelism},
which is a property of an algorithm, and hardware {\em concurrency}, which is a
property of a hardware architecture. Parallelism in an algorithm provides the
opportunity to perform computations concurrently on hardware.  Algorithms often
exhibit multiple levels of parallelism, and hardware often provides multiple
levels of concurrency.  In these terms, the primary contribution of this paper
will be to determine the optimal mappings between the parallelism available in
$\ell_1$-min and the concurrency available in multi-core CPU and GPU
architectures.

Finally, we present extensive benchmark to compare the performance of the
generic ALM $\ell_1$-min algorithm and the corresponding face recognition
routines on massively parallel CPUs and GPUs.  To this end, a new face
recognition system has been fully implemented on a state-of-the-art
workstation for security and access control applications, as shown in Figure \ref{fig:pipeline}. 
Since the parallelization of any algorithm
is highly dependent on the target hardware, we will give a brief overview of
the architecture of our test system in detail in Section
\ref{sec:parallelism}.
Finally, note that for applications targeting thousands of users or more, it would likely be
necessary to distribute the work over a network of computers, while this paper
focuses on the (shared memory) parallelism available in a single machine.
The C++ source code of our implementations, which target Nvidia's CUDA GPU architecture,
will be made available for peer evaluation.
\begin{figure}
\centering
{\tiny \input{figures/pipeline_simplified.tex}}
\caption{The face recognition pipeline.}
\label{fig:pipeline}
\end{figure}

\subsection{Literature Review} 
We first briefly review the literature of $\ell_1$-min and existing
works on its parallelization.  Traditionally, $\ell_1$-min (or the so-called
basis pursuit (BP)) has been formulated as a linear program
\cite{ChenS2001-SIAM}. Several variations of the solution are also well known
in optimization, including a noisy approximation via quadratic programming
called the LASSO \cite{TibshiraniR1996} and truncated Newton interior-point
method (TNIPM) \cite{KimS2007}.

One of the drawbacks of most interior-point methods for $\ell_1$-min (including
BP, LASSO, and TNIPM) is that they require the solution sequence to follow an
interior path via gradient descent or conjugate gradient methods, which are computationally expensive.
To mitigate these issues, an approach called \emph{Homotopy} has been recently studied to accelerate the
speed of $\ell_1$-min. Homotopy methods for recovering sparse signals were
first studied in the context of LASSO \cite{OsborneM2000}, which inspired a
solution to the \emph{forward stagewise linear regression} problem called LARS
\cite{EfronB2004}, and eventually led to the corresponding Homotopy algorithms
for BP in \cite{MalioutovD2005,DonohoD2006}.

Although Homotopy can be shown to exactly estimate BP when the solution is
sufficiently sparse \cite{DonohoD2006}, the algorithm still involves 
computationally expensive operations such as matrix-matrix multiplication and 
the solution of least-squares problems with varying $A$ matrices. 
In Section \ref{sec:ALM}, we contend
that ALM is a better choice among existing $\ell_1$-min
solvers for implementation on many-core CPUs and GPUs. The ALM algorithm belongs to a
category of approximate $\ell_1$-min solutions called \emph{iterative
shrinkage-thresholding} (IST) methods \cite{WrightS2008,BeckA2009}.
IST algorithms mainly involve elementary operations such as vector 
algebra and matrix-vector multiplication. Therefore,
when the dimension of the problem becomes high, IST-type algorithms are
particularly suitable for systems with parallel-computing pipelines. In
\cite{YangA2010-ICIP}, the authors further showed that, compared to other
IST-type algorithms, ALM is able to achieve good estimation accuracy comparable
to BP and Homotopy, while at the same time significantly improve the speed
compared to these exact $\ell_1$-min solutions. Due to these facts, in this
paper, we choose ALM as the core algorithm to implement the complete parallel
face recognition pipeline.

In terms of the past works in parallel $\ell_1$-min, the literature has been
limited, to the best of our knowledge. In \cite{BorghiA2010}, Borghi et al.
developed a special proximal gradient $\ell_1$-min algorithm based on
Moreau-Yosida regularization. In \cite{MurphyM2010}, Murphy et al. presented
parallel implementation of the l1-SPIRIT MRI reconstruction algorithm on NVidia
programmable GPUs, which is similar to the hardware setup of this paper.

\section{Augmented Lagrange Multiplier Method}
\label{sec:ALM}

For the completeness of the paper, in this section, we briefly describe the ALM
algorithm for $\ell_1$-min \eqref{eq:l1min} \cite{YangA2010-ICIP} and analyze
its complexity. Parallelization of ALM for face alignment and recognition will
be discussed in Section \ref{sec:alignment} and Section \ref{sec:recognition},
respectively.

Lagrange multiplier methods have been frequently used in convex programming to
eliminate equality constraints via adding a significant penalty term to the
cost function for infeasible points. ALM methods differ from other
penalty-based approaches by simultaneously estimating the optimal solution and
Lagrange multipliers in an iterative fashion.  For $\ell_1$-min, the augmented
Lagrangian is defined as: \begin{equation} L_\mu(\xx,\yy) = \|\xx\|_1 +
\yy^T(\bb - A\xx - \ee) + \frac{\mu}{2} \| \bb-A\xx-\ee \|_2^2, \end{equation}
where $\mu > 0$ is a constant that penalizes infeasibility and $\yy$ is a
vector of lagrange multipliers.

In Lagrange Multiplier Theory, if there exists a Lagrangian $y^*$ that
satisfies the second-order sufficiency conditions for optimality
\cite{BertsekasD2003}, then for a sufficiently large $\mu$, the optimal
$\ell_1$-min solution also minimizes
\begin{equation}
\xx^* = \arg \min L_\mu(\xx,\yy^*).
\label{eq:optimal-ALM}
\end{equation}

In practice, the optimal values for the triplet $(\xx^*, \yy^*, \mu)$ are all
unknown. Furthermore, it has been observed that solving
\eqref{eq:optimal-ALM} with a large initial value of $\mu$ tends to lead to
slower convergence speed \cite{WrightS2008,YangA2010-ICIP}. In
\cite{BertsekasD2003,YangJ2009}, an alternating procedure has been shown to
iteratively update $\xx$ and $\yy$:
\begin{equation}
\left \{
\begin{array}{lll}
\xx_{k+1} & = & \arg\min_{\xx} \, L_{\mu_k} (\xx,\yy_k)\\
\yy_{k+1} & = & \yy_k + \mu_k (\bb - A\xx_{k+1}) \\
\end{array}
\right . ,
\label{eq:alm}
\end{equation}
where $\mu_{k}\rightarrow \infty$ is a monotonically increasing positive sequence.
The iteration terminates when the estimates $(\xx, \yy)$ converge.

Note that in the iterative procedure \eqref{eq:alm}, the second
step only involves vector algebra and matrix-vector multiplication. Therefore,
the procedure is computationally efficient only if it is easier to minimize the
augmented Lagrangian $L_{\mu_k} (\xx,\yy_k)$ compared to solving the original problem
\eqref{eq:l1min} directly. In fact, this problem can be solved element-wise
iteratively by a soft-thresholding algorithm \cite{WrightS2008,BeckA2009},
whose time complexity is bounded by $O(n^2)$ and can be easily parallelized.
Algorithm \ref{alg:alm} summarizes the generic ALM $\ell_1$-min algorithm.\footnote{For conciseness, we
only present the ALM algorithm in the primal domain. There also
exist implementations in the dual domain \cite{YangJ2009,YangA2010-ICIP}.}

 \begin{algorithm}[h]
\caption{Augmented Lagrangian Method (ALM)}
{\bf INPUT:} $\bb \in \Re^m$, $A=[A_1,\cdots, A_K] \in \Re^{m \times n}$, $\tau\leftarrow \max\mbox{eig}(A^TA)$, and constant $\rho>1$.
\begin{algorithmic}[1]
\WHILE{not converged ($k = 1,2,\ldots$)} 
\STATE $t_1 \leftarrow 1$, $\zz_1 \leftarrow \xx_k$, $\uu_1 \leftarrow \xx_k$ 
\WHILE{not converged ($l = 1,2,\ldots$)} 
\STATE $\uu_{l+1}  \leftarrow \mbox{shrink}(\zz_l - \frac{1}{\tau}A^T(A\zz_l - \bb - \frac{1}{\mu_k}\yy_k), \frac{1}{\mu_k\tau})$
\STATE $t_{l+1} \leftarrow \frac{1}{2}( 1 + \sqrt{1+4t_l^2})$
\STATE $\zz_{l+1} \leftarrow \uu_{l+1}+ \frac{t_l - 1}{t_{l+1}}(\uu_{l+1} - \uu_l)$ 
\ENDWHILE 
\STATE $\xx_{k+1} \leftarrow \uu_{l+1}$ 
\STATE $\yy_{k+1} \leftarrow \yy_k + \mu_k (\bb - A\xx_{k+1})$ 
\STATE $\mu_{k+1} \leftarrow \rho\cdot\mu_k$ 
\ENDWHILE 
\end{algorithmic}

{\bf OUTPUT:} $\xx^* \leftarrow \xx_k$.
\label{alg:alm}
\end{algorithm}

\section{Hardware Concurrency} \label{sec:parallelism}

In this section we discuss the levels of concurrency available in the hardware
architectures considered in this paper, as well as other aspects of the
hardware that are important for performance.  In particular, since on-chip
memory (cache) is often orders of magnitude faster than off-chip memory, the
size and speed of on-chip memory (cache) have big performance implications.
We give a brief overview of the caches that are available in our target
architectures, and defer discussion of their performance implications to
sections \ref{sec:alignment} and \ref{sec:recognition}.

We will primarily compare CPU and GPU architectures on a per-board basis:
for CPU implementation we make use of all the cores in as many CPUs are present
on one motherboard, and for GPU implementation we make use of a single
many-core Nvidia GPU chip on a PCI card. GPUs are commonly shipped with a
single GPU chip per PCI-express card.  Note that most engineering  workstations
are capable of supporting several GPU cards.  Unless otherwise specified, all
implementations utilize single precision floating point datatypes.

\subsection{CPU Hardware Concurrency} 
\label{sec:CPU-concurrency}
The main defining characteristics of contemporary multi-core CPU architectures
are that they have two levels of concurrency, relatively large amounts of
on-chip cache, and relatively high clock speed.

The baseline architecture for our experiment is a Linux workstation with two
quad-core Intel Nehalem E5530 processors clocked at 2.4 GHz.  Each processor
has its own memory interface, and is directly interfaced to half of the RAM
installed in the machine.  The amount of RAM installed exceeds the amount used
by the algorithms, and is not an important performance consideration.

% CPU CACHE DISCUSSION
%2MB = 2**21 bytes
%64*64*4 = 2**14 bytes = 16384 bytes per image
%... so can fit 2**7  = 128 images.
Each core has a private 32\,KB L1 data cache and a private 256\,KB L2 cache (\ie, a total of 2\,MB L2 cache available for the eight cores).
Therefore, for the low-resolution images used in face recognition (say, on the order of $64\times64$ grayscale images), the L2 cache
in principle can hold up to 16 images per core and 128 images in total.
%16MB = 2**24 bytes = 16777216 bytes
%64*64*4 = 2**14 bytes = 16384 bytes per image
%... so can fit 2**10  = 1024 images.
Each processor further has 8\,MB of L3 cache that is shared by the four cores.
Overall, the algorithm has approximately 16\,MB of L3 cache
available for a dual-processor configuration, which corresponds to about 1024 images.

%With Hyperthreading enabled, each core can support two threads simultaneously.
For floating-point instructions, each core also has a vector processing unit
(SSE) capable of performing the same arithmetic operation on four single-precision floating point values simultaneously.

There are thus two important levels of concurrency that need to be exploited to
efficiently use a modern CPU: {\em core-level} concurrency and {\em vector-level}
concurrency.

\subsection{GPU Hardware Concurrency}
The main defining characteristics of contemporary multi-core GPU architectures
are that they have two (much wider) levels of concurrency, relatively small amounts of
on-chip cache, and relatively low clock speed.

Compared to CPU architectures, modern GPU architectures dedicate less of their transistors
to cache, and more of them for arithmetic logic. Our target GPU architecture
is the Nvidia Fermi architecture (\eg, GTX 480 GPU used in this paper). CUDA based GPUs are comprised of several 
\emph{streaming multiprocessors} (SMP) (currently 14 to 16 for a high end GPU).

The GPU executes programs as SIMT (Single Instruction Multiple Threads) in
groups of 32 threads called \emph{warps}.  These warps are then grouped into
higher levels called \emph{thread blocks} and \emph{grids}.  Since SMPs are not designed
to communicate between themselves during GPU program execution, tasks
are typically broken up into multiple GPU programs (kernels) that are executed
serially.\footnote{Please add a footnote to further explain.}

Each SMP contains its own L1 cache, which is divided between hardware managed
and software managed (``shared'') memory.  Additionally, all SMPs share a common
L2 cache.  For our system, each SMP has 64 KB of L1 cache with a total of 896
KB of L1 cache/shared memory and 768KB of L2 cache. The relatively small amount
of cache (1/23 as much as CPU L3) on the GPU is ballanced by a significantly
higher bandwidth between the processor chip and off-chip memory (DRAM) compared
to the CPU.

Compared to typical CPUs, GPUs have much higher peak computational bandwidth
as well as data bandwidth to DRAM.  However, since our system primarily consists
of matrix-vector operations and image resampling operations, it has a very low
arithmetic intensity (\ie, we perform a very small number of arithmetic
operations for every number that has to be loaded), and thus performance is
typically constrained by memory (and cache) bandwidth.  At large problem sizes
the CPU is limited by its relatively low bandwidth to DRAM compared to the GPU,
while at small problem sizes the GPU is limited by the relatively high latency
of its access to DRAM.

\section{Parallelism in the Face Alignment Stage}
\label{sec:alignment}

Given the hardware concurrency in modern parallel computing architectures, next,
we shall study the optimal mappings of the algorithm parallelism in face recognition to the multi-core CPUs and GPUs, respectively.
In this section, we first focus on the face alignment stage (see Figure \ref{fig:pipeline}).

Face alignment \eqref{eq:l1min_alignment} estimates an image transformation $\tau$ that rectifies the query image $\tilde{\bb}$ with possible pose variation
w.r.t. each training class $A_i$, which leads to the minimal sparsity in error $\ee$ after the alignment. Note that directly solving \eqref{eq:l1min_alignment} would require
simultaneously estimating $\xx$, $\ee$, and $\tau$ typically on a nonlinear manifold, which renders the problem extremely difficult to solve exactly with many local minima.

To mitigate the problem, if the system can obtain to a good initial guess of the transformation $\tau$ (\eg, provided by a good face detector),
the optimal solution for $\tau$ can be sought iteratively by a sequence of linearized approximations $j=1, 2, \cdots$ as follows:
\begin{equation}
\min_{\xx, \ee, \Delta \tau_j}\|\ee\|_1\quad \subj\quad \tilde\bb\circ\tau_j +  J_j\Delta \tau_j = A_i\xx + \ee,
\end{equation}
where $J_j = \nabla_{\tau_j}(\tilde\bb\circ\tau_j)$ is the Jacobian, and $\Delta \tau$ is a step update to $\tau$. Denote $\bb_j = \tilde\bb\circ\tau_j$, $B_j = [A_i, -J_j]$ and $\ww^T = [\xx^T, \Delta \tau^T]$, then the update $\Delta \tau$ can be computed by solving the following linear program:
\begin{equation}
\min_{\ww, \ee}\|\ee\|_1\quad \subj\quad \bb_j = B_j\ww + \ee.
\label{eq:alignment-linearization}
\end{equation}

The per-class alignment algorithm via ALM is summarized in Algorithm \ref{alg:alignment_alm}. In terms of computational complexity, one significant term is the computation of the matrix psudo-inverse $B_j^\dagger$, whose computational cost is bounded by $O(n_i^2m + n_i^3)$. For per-class face alignment, since both $A_i$ and $J_{j}$ are very tall matrices (\ie, $m\gg n_i$ for all $i=1, \cdots, K$ classes), we find that the computation time is still dominated by the iterative computation of matrix-vector multiplications in the inner loop.
\begin{algorithm}[ht!]
\caption{\bf (Face Alignment via ALM)} \label{alg:alignment_alm}
{\bf Input:} $\bb$, $A_i$, $\x_0 = \mathbf{0}$, $\tau_0$, and $J_0$.
\begin{algorithmic}[1]
\WHILE{not converged ($j = 1,2,\ldots$)}
\STATE Update $\bb_j \leftarrow \frac{\bb\circ \tau_{j-1}}{\|\bb\circ \tau_{j-1}\|}$; $B_j= [A_i, -J_{j-1}]$ and corresponding $(B_j^\dagger)^T$
\STATE Initialize $\ww_0 = \mathbf 0$, $\yy_0 = \mathbf 0$
\WHILE{not converged ($k = 1,2,\ldots$)}
\STATE $\uu_0\leftarrow \ww_{k-1}$; $\zz_0\leftarrow \ee_{k-1}$
\WHILE{not converged ($l = 1,2,\ldots$)}
\STATE $\zz_l \leftarrow \mbox{shrink}\left(\bb_j - B_j\uu_{l-1} + \frac{\yy_{k}}{\mu_{k-1}}, \frac{1}{\mu_{k-1}}\right)$
\STATE $\uu_l \leftarrow (B_j^\dagger)^T \left(\bb_j - \zz_{l} + \frac{\yy_{k-1}}{\mu_{k-1}} \right) $
\ENDWHILE
\STATE $\ww_k \leftarrow \uu_l$; $\ee_k \leftarrow \zz_l$
\STATE $\yy_{k} \leftarrow \yy_{k-1} + \mu_{k-1} (\bb_j - B_j\ww_{k} - \e_{k})$
\STATE $\mu_{k} \leftarrow \rho\mu_{k-1}$
\ENDWHILE
\STATE Update $\ee_j$, $\tau_j$, and $J_j$
\ENDWHILE
\end{algorithmic}
{\bf Output:} $\tau^*\leftarrow \tau_j, \e^*\leftarrow \ee_j$
\end{algorithm}

In summary, the alignment stage essentially contains two levels of available
parallelism. At the higher level, there are per-class alignment problems
that are solved independently, or \emph{problem-level parallelism}.  At a lower
level, the first-order linear algebraic operations exhibit parallelism within
image operations, \ie, at the pixel level.  We call this \emph{pixel-level
parallelism}.  In the following two sections will will discuss how to
exploit these two levels of parallelism on different CPU and GPU architectures.

\subsection{CPU Implementation} 
\label{sec:alignment_implementation_cpu}

The optimal strategy to implement Algorithm \ref{alg:alignment_alm} on a multicore CPU must take into account
the trade-off between the available memory/cache sizes at multiple hierarchies and their bandwidth based on a particular
CPU architecture. Figure \ref{fig:alignment_mapping_old} shows the the sizes of L2 and L3 caches of the Intel E5530 CPU, 
which is a good example of modern multicore CPU architectures. 
\begin{figure}
\centering
\includegraphics[width=3.4in]{figures/alignment_mapping_old}
\caption{Available L2 and L3 caches on the Intel E5530 multicore CPU architecture with dual CPU configuration.}
\label{fig:alignment_mapping_old}
\end{figure}

As we have calculated in Section \ref{sec:CPU-concurrency}, the L2 cache in each core is only able to store about 16 images. For image alignment problem, the number of training images per class is typically larger than 16. Hence, if \eqref{eq:alignment-linearization} would be solved in parallel on the eight cores, it would require L3 cache to hold data up to four classes at a time, and the L2 cache should retrieve subsets of the data frequently during the iterative process to execute Algorithm \ref{alg:alignment_alm}. Therefore, the bottleneck in this strategy becomes the bandwidth between the L2 cache and L3 cache. The implementation of some previous works \cite{WagnerA2009-CVPR,WagnerA2011-PAMI} is indeed based on this mapping strategy.

In this paper, we show that for CPU parallelism, a more traditional implementation for pixel-level parallelism in fact outperforms problem-level parallelism using eight separate cores. In other words, the CPU implementation performs alignment of a query image against the training images of single or dual users at a time on the dual processors, depending on the setting of the compiler. More specifically, in our implementation, most of the operations take advantage of Intel Math Kernel Library (MKL), a commercial implementation of the standard Basic Linear Algebra Subprograms (BLAS). For the image resampling step, we make use of the Intel Integrated Performance Primitives (IPP) library. Both MKL and IPP are optimized for Intel multicore CPUs, and are able to automatically utilize both core-level and vector-level concurrency. For other operations that are not optimized by Intel in-house libraries, we achieve both vector-level and core-level concurrency through a combination of the Open MP API \cite{dagum2002openmp} and the automatic vectorization facilities of the Intel C++ compiler (ICC) \cite{dulong1999overview}. We shall compare the performance of the two parallelism methods in Section \ref{sec:experiment}.

\subsection{GPU Implementation}
\label{sec:alignment_implementation_gpu}

Similarly, we also consider the trade-off between the available cache sizes and their bandwidth based on the Nvidia CUDA GPU architecture chosen for the paper. Figure \ref{fig:alignment_mapping_gpu} illustrates the multicore architecture and the L1/L2 caches of the Nvidia GTX 480 GPU.
\begin{figure}
\centering
\includegraphics[width=3.4in]{figures/alignment_mapping_gpu}
\caption{Proposed mapping of alignment parallelism onto gpu concurrency}
\label{fig:alignment_mapping_gpu}
\end{figure}

Our proposed GPU implementation for the iterative alignment stage consists of a
single kernel that performs all of the test image alignment problems.  Each of
instance of the alignment problem is assigned its own thread block, and the GPU
hardware schedules as many of these thread blocks to run concurrently as the 
hardware will allow.

In addition, the full resolution query image, along with its numerical derivatives in both
image directions, are stored in GPU texture memory. This  allows the
computation of $b(\tau)$ and $J(\tau)$, which essentially consist of warped
versions of the test image and its derivatives, to exploit the GPU's special
purpose hardware for bilinear interpolation.

In one experiment, we have exploited several possible GPU implementation strategies, which include
solving single alignment problems one at a time using CUBLAS (302\,ms); solving one problem
per SMP using CUDA streams (70\,ms); and solving one problem per SMP using one large kernel call
 (36\,ms). The inferior performance of CUBLAS may be attributed to the relatively small problem
 size. In Section \ref{sec:experiment}, we will further exploit the difference between CPU and GPU
 via synthetic and real image data.

\section{Parallelization of the Face Recognition Stage}
\label{sec:recognition}
After the face alignment stage is complete, a single recognition algorithm will be called
to estimate a sparse representation of the query image w.r.t. all the training classes, 
as shown in \eqref{eq:l1min_denoise}. In \cite{WrightJ2009-PAMI,YangA2010-ICIP,WagnerA2011-PAMI},
the recognition algorithm via $\ell_1$-min has been extensively studied. The algorithm itself via ALM
is summarized in Algorithm \ref{alg:alm_rec}.
In this section, we mainly focus our discussion on its pixel-level parallelism on the multicore CPUs and GPUs. Then in Section \ref{sec:experiment}, we benchmark
the performance of the two architectures and further demonstrate the improvement in both accuracy and speed
compared to the existing literature.
\begin{algorithm}[t]
\caption{\bf (Face Recognition via ALM)} \label{alg:alm_rec} 
\begin{algorithmic}[1]
\begin{small}
\STATE {\bf Input:} $\bb \in \Re^m$, $A \in \Re^{m \times n}$,
$\x_1 = \mathbf{0}$, $\e_1 = \bb$, $\y_1 =
\mathbf{0}$.
\WHILE{not converged ($k = 1,2,\ldots$)}
\STATE $\e_{k+1} = \textup{shrink}(\bb - A\x_k +
\frac{1}{\mu}\y_k, \frac{1}{\mu})$;
\STATE $t_1\leftarrow 1$, $\z_1 \leftarrow \x_k$, $\w_1 \leftarrow \x_k$;
\WHILE{not converged ($l = 1,2,\ldots$)}
\STATE $\w_{l+1} \leftarrow \textup{shrink}(\z_l +
\frac{1}{\gamma}A^T(\bb - A\z_l - \e_{k+1} +
\frac{1}{\mu}\y_k), \frac{1}{\mu\gamma})$;
\STATE $t_{l+1} \leftarrow \frac{1}{2}( 1 +
\sqrt{1+4t_l^2})$;
\STATE $\z_{l+1} \leftarrow \w_{l+1} + \frac{t_l - 1}{t_{l+1}}(\w_{l+1} - \w_l)$;
\ENDWHILE
\STATE $\x_{k+1} \leftarrow \w_{l}$,  \; $\y_{k+1} \leftarrow \y_k + \mu (\bb - A\x_{k+1} - \e_{k+1})$;
\ENDWHILE \STATE
{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
\end{small}
\end{algorithmic}
\end{algorithm}

For CPU implementation, {\bf AY: Drew, please fill in one or two paragraphs}.

For GPU implementation, {\bf AY: Drew, please fill in one or two paragraphs}.

\section{Experiment}
\label{sec:experiment}
In this section, we benchmark the performance of our parallel implementations
of $\ell_1$ minimization on CPU and GPU platforms.  In order to show how our
$\ell_1$ minimization algorithms scale with problem size, in Section
\ref{sec:simulation} we will begin with benchmarks on synthetic data.  We will
then progress in Sections \ref{sec:alignment_benchmark} and
\ref{sec:recognition_benchmark} to demonstrating the speed of our
implementations as applied to the alignment and recognition stages of the face
recognition problem.

To generate a meaningful comparison between different hardware architectures,
in all cases we compare the performance of hardware architectures on a
per-board basis:  For CPU implementations, the benchmark makes use of all of
the cores in as many cpu's are present.  For GPU implementations, the benchmark
makes the best use of the entire GPU chip (most GPU boards have a single GPU
chip).  Unless otherwise specified, all implementations utilize single-precision floating point datatypes.

\subsection{Simulations on Random Data}
\label{sec:simulation}

The first experiment compares the performance of our two proposed implementations of 
the general $\ell_1$-min (Algorithm \eqref{alg:alm}) given a generic random matrix $A$ and a random sparse
vector $\xx_0$.  
The measurement $d \times n$ matrix $A$ was generated as a random Gaussian matrix, with each entry
randomly generated from the standard normal and normalized with unit column norm.  The ratio of $d:n$ varied from
$1:10$ to $5:10$ with $n$ varying from 500 to 8000.
The ground truth signals, $\xx_0$ had a sparsity rate of 0.1 with elements generated using the standard normal and also normalized with unit norm.
Because the ground truth signal $\xx_0$ was known, the algorithm terminated
when $\|\bx-\bx_0\| < \tau$.
The measurement vector was generated with the equation $\bb_0 = A \xx_0$.

The results of this benchmark can be found in Figure \ref{fig:random_data}.
The x-axis represents the size of the $A$ matrix and the y-axis represents the
average amount of time it to complete a single problem instance.  The GPU implementation tends to be faster than
the CPU implementation at solving a single large problem, whereas the the CPU
implementation is faster at solving a single small problem.  Figure \ref{fig:random_data2} shows the various sizes and tolerances at which our GPU implementation
begins to run faster than the CPU implementation on synthetic data and will provide researchers 
with a starting point of where to begin setting their tolerance levels for the ALM algorithm.
\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{results/random_data/time_vs_matrix_size_constant_tol}
\end{center}
\caption{Speed vs Dimensions of $A$.}
\label{fig:random_data}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{results/random_data/size_vs_speed_crossover_ratio_3}
\end{center}
\caption{Relation between CPU and GPU runtime vs tolerance level}
\label{fig:random_data2}
\end{figure}

\subsection{Face Alignment Stage Benchmark} 
\label{sec:alignment_benchmark}
This section presents benchmarks of our CPU and GPU implementations of the
alignment stage of the face recognition pipeline described in Algorithm
\ref{alg:iterative_alignment}.  The iterative alignment seeks a similarity
transformation between the coordinate frame of the "camera" resolution test
image and the coordinate frame of the "canonical" resolution images which the
images are compared.  We use a $64$ pixel wide square window in the canonical
frame.  The training images are aligned by manually clicking the two outer eye
corners in the camera frame, and then warping the images via a similarity
transformation such that the eye corners lie at a pair of canonical eye
locations in the canonical frame.  The canonical eye corners are chosen such
that for well aligned images, the canonical window contains exclusively pixels
from the subject's face.

For experiments on face data we use subsets of the CMU Multi-PIE Face Database.
For gallery images we use frontal images from session 1, which contains 20
images per user taken under different illuminations. For test images we use
images from session 2.

In Figure \ref{fig:alignment_stage_runtime} we report the time it takes
to align a single test image against each of the users in the gallery.
We vary the size of the gallery to show how implementation scales in the
number of gallery users.
\begin{figure}
\centering
\includegraphics[width=3.4in]{figures/alignment_runtime_graph}
\caption{Alignment Stage Runtime vs. Size of Training Database}
\label{fig:alignment_stage_runtime}
\end{figure}

Previous publications used a hard threshold on the number of subjects kept for the recognition
stage as a compromise between recognition speed and potential recognition rate.
In order to better motivate 
a choice for a threshold of the number of users kept for the recognition stage,
we varied the number of users kept for the recognition stage and estimated the probability
of the test user being among the set of training users kept for the recognition stage.
As can be seen in Figure \ref{fig:user_alignment_rank_plot}, there are rapidly diminishing
returns as the number of kept users is increased beyond ten.  
\begin{figure}
\centering
\includegraphics[width=3.4in]{figures/user_alignment_rank_plot}
\caption{Probability of the test user being kept for the recognition stage vs. Number of traning
users kept for the recognition stage}
\label{fig:user_alignment_rank_plot}
\end{figure}

\subsection{Face Recognition Stage Benchmarks}
\label{sec:recognition_benchmark}

\subsection{Face Recognition Pipeline Benchmarks}
\label{sec:pipeline_benchmark}
The fourth experiment compares the recognition rate of the face recognition algorithm at different
image resolutions.  We fixed the number of users to be $enter here$ which was shown in figure  $enter here$  have 95\% probability
to have the true user in the training set, inner loop tolerance to 1e-2 and outer loop tolerance to 1e-3.

Our final experiment is to compare the speeds of CPU and GPU implementations entire pipeline in Figure \ref{fig:pipeline}.

\section{Conclusion}

{\small
\bibliographystyle{ieee}
\bibliography{faces}
}

\end{document}